Understanding the Socio-Economic Determinants of Voter Behavior: An Analysis of the 2020 U.S. Presidential Election

### 2024-03-16 paper/us_election_2020.qmd
final edits

We expect a positive relationship between individuals of Black, Asian, and Hispanic ethnicities and their support for Biden. This expectation arises from Trump's history of spreading polarizing language and anti-immigrant sentiments, as well as his controversial plans such as building a border wall [@trump_wall]. These groups are more likely to align with Biden's policies, which prioritize inclusion and diversity [@biden_inclusion]. White individuals with traditional family values and conservative leanings tend to support Trump. They are drawn to his emphasis on preserving traditional values and promises to uphold conservative principles, especially regarding immigration, law and order, and gun rights.

Conversely, we anticipate a negative relationship between voters in the South and Midwest regions and support for Biden. These regions have a stronger conservative presence and a history of supporting Republican candidates like Trump. States, such as Texas, which are known Republican strongholds, are located in the South. Therefore, individuals in these regions may be less inclined to support Biden's progressive agenda.

Regarding employment status, we expect students, unemployed individuals, and those temporarily laid off to be more inclined to support Biden. Students are often exposed to diverse perspectives and progressive ideas in educational settings, making them more likely to endorse Biden's platform. Unemployed and laid-off individuals may favor Biden due to the Democratic Party's advocacy for social welfare programs and support for workers' rights.

@@

The reason for the majority of Hispanics supporting Biden may be due to the US-Mexico border wall that was built during Trump's reign. The wall was built over 452 miles, which replaced the old fence [@Rodriguez_2021]. With this measure in place, Trump lost most of the Hispanic votes. Trump claimed the wall reformed the immigration system and achieved the most secure southern border in US history. However, that was not true, and with it had come a surge in illegal border crossings [@trump_wall], which overall made Hispanic people dissatisfied with Trump's decisions and ruling.

As COVID-19 pandemics spreaded, Trump repeatedly called it "Chinese virus" [@trump_virus]. This obviously led to a large number of Asians, especially Chinese people, feel targeted becuase it is an extremely anti-Asian statement: although COVID-19 was first found in China, but it does not mean that virus was from China. Trump's comments quickly festered in public opinion and led to a dramatic increase in hate incidents in the Asian American community [@Reja_2021]. The repeated use of the "Chinese Virus" by a president of the country has contributed to an environment of racial discrimination and hatred against Asians. Therefore, they became dissatisfied with Trump and more supportive to Biden, whose policies are more open and democratic, and values racial equality more.

@@

Our analysis highlights regional voting variations, with the Northeast, West, and Midwest favoring Biden, while the South supports Trump. These differences stem from diverse cultural, family, religious, and social perspectives.

Biden's support base is concentrated in the Mid-Atlantic states, New England, the West Coast, Southwestern states, and certain regions around the Great Lakes [@wiki2020]. This support is particularly evident in cities within these regions, driven by various factors including liberal ideals and progressive policies. For instance, states like New York and Massachusetts support Biden due to their liberal-leaning populations, diverse demographics, and emphasis on social justices. Similarly, on the West Coast, states such as California and Washington are known for their progressive values and advocacy for issues like environmental conservation and LGBTQ+ rights, contributing to Biden's popularity in these areas. In the Southwest, states like Arizona and Nevada, with growing minority populations and concerns over immigration policies, also show significant support for Biden. Additionally, states around the Great Lakes, including Illinois and Michigan, align with Biden's platform due to their industrial heritage and strong labor unions.

@@

>>>>

We expect a positive relationship between individuals of Black, Asian, and Hispanic ethnicities and their support for Biden. This expectation arises from Trump's history of spreading polarizing language and anti-immigrant sentiments, as well as his controversial plans such as building a border wall [@Rodriguez_2021]. These groups are more likely to align with Biden's policies, which prioritize inclusion and diversity [@biden_inclusion]. White individuals with traditional family values and conservative leanings tend to support Trump. They are drawn to his emphasis on preserving traditional values and promises to uphold conservative principles, especially regarding immigration, law and order, and gun rights.

Conversely, we anticipate a negative relationship between voters in the South and Midwest regions and support for Biden. These regions have a stronger conservative presence and a history of supporting Republican candidates like Trump [@wiki_south]. States, such as Texas, which are known Republican strongholds, are located in the South. Therefore, individuals in these regions may be less inclined to support Biden's progressive agenda.

Regarding employment status, we expect students, unemployed individuals, and those temporarily laid off to be more inclined to support Biden. Students are often exposed to diverse perspectives and progressive ideas in educational settings, making them more likely to endorse Biden's platform. Unemployed and laid-off individuals may favor Biden due to the Democratic Party's advocacy for social welfare programs and support for workers' rights [@wiki_biden].

@@

The reason for the majority of Hispanics supporting Biden may be due to the US-Mexico border wall that was built during Trump's reign. The wall was built over 452 miles, which replaced the old fence [@Rodriguez_2021]. With this measure in place, Trump lost most of the Hispanic votes. Trump claimed the wall reformed the immigration system and achieved the most secure southern border in US history. However, that was not true, and with it had come a surge in illegal border crossings [@Rodriguez_2021], which overall made Hispanic people dissatisfied with Trump's decisions and ruling.

As COVID-19 pandemics spreaded, Trump repeatedly called it "Chinese virus" [@trump_virus]. This obviously led to a large number of Asians, especially Chinese people, feel targeted because it is an extremely anti-Asian statement: although COVID-19 was first found in China, but it does not mean that virus was from China. The repeated use of the "Chinese Virus" by a president of the country quickly festered in public opinion and led to a dramatic increase in racial discrimination and hate incidents towards Asian American communities [@Reja_2021]. Therefore, they became dissatisfied with Trump and more supportive to Biden, whose policies are more open and democratic, and values racial equality more [@wiki_biden].

@@

Our analysis highlights regional voting variations, with the Northeast, West, and Midwest favoring Biden, while the South supports Trump. These differences stem from diverse cultural, religious, and social perspectives based on regions.

Biden's support base is concentrated in the Mid-Atlantic states, New England, the West Coast, Southwestern states, and certain regions around the Great Lakes [@wiki2020]. This support is particularly evident in cities within these regions, driven by various factors including liberal ideals and progressive policies. For instance, states like New York and Massachusetts support Biden due to their liberal-leaning populations, diverse demographics, and emphasis on social justices. Similarly, on the West Coast, states such as California and Washington are known for their progressive values and advocacy for issues like environmental conservation and LGBTQ+ rights[@lgbtq], contributing to Biden's popularity in these areas. In the Southwest, states like Arizona and Nevada, with growing minority populations and concerns over immigration policies, also show significant support for Biden. Additionally, states around the Great Lakes, including Illinois and Michigan, align with Biden's platform due to their industrial heritage and strong labor unions.

@@

It is evident that Trump gained more support from homemakers, retirees, and permanently disabled individuals. This preference could be due to the fact that these groups may not be directly impacted by the economic policies proposed by either candidate. Thus, their voting decisions may be influenced by other campaign aspects. Biden's focus on creating job opportunities and raising minimum wages may strongly resonate with those who are actively working or seeking employment [@biden_inclusion], but it may not hold as much relevance for homemakers, retirees, and the permanently disabled.
